---
title: "Misc NVS data preparation"
# subtitle: 
# author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
mainfont: DejaVu Sans
output: 
  html_document: 
    highlight: pygments
    keep_md: no
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width=9, fig.height=7, dpi=300, out.width="900px", out.height="700px")

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

options(scipen=999)
set.seed(12345)

library(pacman) 
p_load(tidyverse, magrittr, readr, readxl, lubridate, 
       scales, janitor, kableExtra, 
       visdat, mvtsplot, 
       sf, tmap, tmaptools)
tmap_mode("view")
```

```{r}
SA3 <- readRDS("Z:/AIRBNB-Q0931/TEMPO_Airbnb/data/geo/SA3_2016_AUST_clean.rds") %>% 
  select(-AREASQKM16, -starts_with("SA4"), -starts_with("GCC"))

SA2 <- readRDS("Z:/AIRBNB-Q0931/TEMPO_Airbnb/data/geo/SA2_2016_AUST_clean.rds") %>% 
  st_drop_geometry() %>% 
  select(SA2_MAIN16, SA2_NAME16, SA3_CODE16, SA3_NAME16)

```

<!-- ------------------------------------------------------------ --> 

# Data

## Origins

Using data from `Home SA2, Month returned from trip and Stopover reason by Nights (000).csv`

Removed observations with region codes like: `Not asked`, `Offshore and Migratory ___`, `Unknown Region ___`, `Transit  ___`,  `Unknown SA2`

`SA3`s from Other Teritories were removed. 

```{r include=FALSE}
origin <- read_csv("data/NVS/raw/Home SA2, Month returned from trip and Stopover reason by Nights (000).zip", skip = 11) %>% 
  clean_names() %>%
  remove_empty(c("rows", "cols")) %>% 
  rename(SA2_NAME16 = home_state_region_sa2,
         month = month_returned_from_trip,
         visitor_nights = x4) %>% 
  mutate(month = factor(month, levels = month.name)) %>% 
  fill(SA2_NAME16, month) %>% 
  filter(SA2_NAME16 != "Total") %>% 
  filter(! SA2_NAME16 %in% c("Not asked")) %>% 
  filter(! grepl('Offshore and Migratory', SA2_NAME16)) %>% 
  filter(! grepl('Unknown Region', SA2_NAME16)) %>% 
  filter(! grepl('Unknown SA2', SA2_NAME16)) %>% 
  filter(! grepl('Transit ', SA2_NAME16)) 

temp <- origin %>% 
  left_join(SA2) %>% 
  filter(is.na(SA3_CODE16)) %>% 
  group_by(SA2_NAME16) %>% 
  summarise(total = sum(visitor_nights, na.rm = TRUE))

origin %<>% 
  left_join(SA2) %>% 
  filter(! is.na(SA3_CODE16)) %>% 
  filter(! SA3_NAME16 %in% c("Christmas Island", "Cocos (Keeling) Islands", 
                             "Jervis Bay", "Norfolk Island")) %>% 
  group_by(SA3_CODE16, SA3_NAME16, stopover_reason, month) %>% 
  summarise(visitor_nights = sum(visitor_nights)) %>% 
  ungroup()
```

There are `r comma(nrow(temp))` locations that do not link to the full list of `SA2_NAME16` codes form ABS. All of them have `nfd` suffix in the name and most have `0` visitors throught the year so most likely these are obsolete codes. These observations have been removed.

```{r}
slice(temp)
```

Median values of guest nights per month per reason:

```{r}
temp <- origin %>% 
  group_by(stopover_reason, month) %>% 
  summarize(median = median(visitor_nights))

temp %>% 
  filter(! stopover_reason %in% c("No other reason", "Not stated/not asked",
                                  "In transit", "Other reason")) %>% 
  ggplot(aes(x = month, y = median, group = stopover_reason, colour = stopover_reason)) + 
  geom_line() +
  theme_light() +
  scale_y_continuous(limits = c(0, 60)) +
  labs(x = "", y = "Median GN", colour = "Stopover reason")
```

## Destinations

Using data from `Stopover SA2, Month returned from trip and Stopover reason by Nights (000).csv`

Removed observations with region codes like: `Not asked`, `Offshore and Migratory ___`, `Unknown Region ___`, `Transit  ___`,  `Unknown SA2`

`SA3`s from Other Teritories were removed. 

```{r include=FALSE}
destination <- read_csv("data/NVS/raw/Stopover SA2, Month returned from trip and Stopover reason by Nights (000).zip", skip = 11) %>% 
  clean_names() %>%
  remove_empty(c("rows", "cols")) %>% 
  rename(SA2_NAME16 = stopover_state_region_sa2,
         month = month_returned_from_trip,
         visitor_nights = x4) %>% 
  mutate(month = factor(month, levels = month.name)) %>% 
  fill(SA2_NAME16, month) %>% 
  filter(SA2_NAME16 != "Total") %>% 
  filter(! SA2_NAME16 %in% c("Not asked")) %>% 
  filter(! grepl('Offshore and Migratory', SA2_NAME16)) %>% 
  filter(! grepl('Unknown Region', SA2_NAME16)) %>% 
  filter(! grepl('Unknown SA2', SA2_NAME16)) %>% 
  filter(! grepl('Transit ', SA2_NAME16)) 

temp <- destination %>% 
  left_join(SA2) %>% 
  filter(is.na(SA3_CODE16)) %>% 
  group_by(SA2_NAME16) %>% 
  summarise(total = sum(visitor_nights, na.rm = TRUE))

destination %<>% 
  left_join(SA2) %>% 
  filter(! is.na(SA3_CODE16)) %>% 
  filter(! SA3_NAME16 %in% c("Christmas Island", "Cocos (Keeling) Islands", 
                             "Jervis Bay", "Norfolk Island")) %>% 
  group_by(SA3_CODE16, SA3_NAME16, stopover_reason, month) %>% 
  summarise(visitor_nights = sum(visitor_nights)) %>% 
  ungroup()
```

Again, there are `r comma(nrow(temp))` locations that do not link to the full list of `SA2_NAME16` codes form ABS. All of them have `nfd` suffix in the name and most have `0` visitors throught the year so most likely these are obsolete codes. These observations have been removed.

```{r}
slice(temp)
```

Median values of guest nights per month per reason:

```{r}
temp <- destination %>% 
  group_by(stopover_reason, month) %>% 
  summarize(median = median(visitor_nights))

temp %>% 
  filter(! stopover_reason %in% c("No other reason", "Not stated/not asked",
                                  "In transit", "Other reason")) %>% 
  ggplot(aes(x = month, y = median, group = stopover_reason, colour = stopover_reason)) + 
  geom_line() +
  scale_y_continuous(limits = c(0, 60)) +
  theme_light() +
  labs(x = "", y = "Median GN", colour = "Stopover reason")

```

```{r eval=FALSE, include=FALSE}
destination %>% 
  filter(stopover_reason %in% c("Holiday")) %>% 
  ggplot(aes(x = month, y = visitor_nights, group = SA3_CODE16)) + 
  geom_line() +
  theme_light() +
  labs(x = "", y = "Median GN", colour = "Stopover reason")
```

```{r eval=FALSE, include=FALSE}
temp <- destination %>% 
  filter(stopover_reason %in% c("Holiday")) %>%
  select(SA3_NAME16, month, visitor_nights) %>% 
  spread(SA3_NAME16, visitor_nights)

mvtsplot(temp, norm = "global")
```

<!-- ------------------------------------------------------------ --> 

# OD relationship

## Reason: holidays

```{r}
max <- max(max(origin$visitor_nights), max(destination$visitor_nights))

rename(origin, origin = visitor_nights) %>% 
  left_join(rename(destination, destination = visitor_nights)) %>% 
  filter(stopover_reason == "Holiday") %>% 
  ggplot(aes(x = origin, y = destination)) +
  geom_line(aes(x = destination, y = destination), colour = "darkorchid4", size = 1) +
  geom_point() +
  theme_light() +
  facet_wrap(vars(month)) +
  coord_fixed(xlim = c(0, max), ylim = c(0, max)) +
  labs(x = "GN origin", y = "GN destination", colour = "Stopover reason")
```

## Reason: business

```{r}
max <- max(max(origin$visitor_nights), max(destination$visitor_nights))

rename(origin, origin =  visitor_nights) %>% 
  left_join(rename(destination, destination =  visitor_nights)) %>% 
  filter(stopover_reason == "Business") %>% 
  ggplot(aes(x = origin, y = destination)) +
  geom_line(aes(x = destination, y = destination), colour = "darkorchid4", size = 1) +
  geom_point() +
  theme_light() +
  facet_wrap(vars(month)) +
  coord_fixed(xlim = c(0, max), ylim = c(0, max)) +
  labs(x = "GN origin", y = "GN destination", colour = "Stopover reason")
```

<!-- ------------------------------------------------------------ --> 

# Net

```{r}
SA3_net <- SA3 %>% 
  left_join(
    rename(origin, origin =  visitor_nights) %>% 
      left_join(rename(destination, destination =  visitor_nights)) %>% 
      mutate(net = origin - destination) %>% 
      select(-origin, -destination))
```

## Jan vs. May, reason: holidays

```{r}
SA3_net %>% 
  filter(stopover_reason == "Holiday") %>% 
  filter(month %in% c("January", "May")) %>% 
  tm_shape() +
  tm_polygons("net", n = 7, palette = "RdYlBu") +
  tm_facets(by = "month")
```

## Jan vs. May, reason: business

```{r}
SA3_net %>% 
  filter(stopover_reason == "Business") %>% 
  filter(month %in% c("January", "May")) %>% 
  tm_shape() +
  tm_polygons("net", n = 7, palette = "RdYlBu") +
  tm_facets(by = "month")
```