---
title: "Facebook data - EDA 03"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
mainfont: DejaVu Sans
output: 
  html_document: 
    # css: custom.css
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    theme: united 
    highlight: pygments 
---

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=8, fig.height=6, dpi=300, out.width="800px", out.height="600px")

options(scipen=999)
set.seed(12345)

library(pacman) 
p_load(tidyverse, readxl, magrittr, anytime, here,
       skimr, sjmisc, naniar, kableExtra, scales, vcd, 
       sf, tmap, tmaptools)

tmap_mode("view") # makes map interactive

```


```{r geodata_prep, eval=FALSE, include=FALSE}
library(rmapshaper)

# postcode outlines 
# exclude 'No usual address (Aust.)' 
# exclude 'Migratory - Offshore - Shipping (Aust.)'

POA <- sf::read_sf(here("data", "geo", "1270055003_poa_2016_aust_shape", "POA_2016_AUST.shp")) %>% 
  filter(POA_CODE16 != "9494") %>% 
  filter(POA_CODE16 != "9797") 

# table(POA$POA_CODE16 == POA$POA_NAME16)
POA$POA_NAME16 <- NULL

class(POA); nrow(POA); names(POA); st_crs(POA); str(POA)
format(object.size(POA), units = "Mb")

# View(POA)
# qtm(POA)

POA_centr <- st_centroid(POA, of_largest_polygon = TRUE)
qtm(POA_centr)
saveRDS(POA_centr, file = here("data", "geo", "POA_2016_AUST_centr.rds"))

POA_simp <- ms_simplify(POA, keep = 0.15, weighting = 0.75) # default settings

format(object.size(POA_simp), units = "Mb")
nrow(POA_simp)
qtm(POA_simp)

# sf::write_sf(POA_simp, here("data", "geo", "POA_2016_AUST_simp.shp"), delete_layer = TRUE)
saveRDS(POA_simp, file = here("data", "geo", "POA_2016_AUST_simp.rds"))
```


```{r geodata_load, include=FALSE}
POA_simp <- readRDS(file = here("data", "geo", "POA_2016_AUST_simp.rds"))
POA_centr <- readRDS(file = here("data", "geo", "POA_2016_AUST_centr.rds"))
```


# Data collection

> Each row represents the response from a GET request

See desc from Alex for further details. 

# Data description

**estimated_dau** – estimated daily average users

**estimated_mau** – estimated monthly average users

Also from Alex:

> Response errors are recorded by the presence of a JSON string in the 'error_message' column; nulls in 'error_message' mean no request/response errors


# Data prep

```{r fbook_prepare, eval=FALSE, include=FALSE}
p_load(bigrquery)

project <- "goopla-1509504704864"
# project <- "twitter-stream-205107"

# sql <- "SELECT * FROM `twitter-stream-205107.twitter_stream.facebook_updated`"
# 
# tb <- bq_project_query(project, sql)
# facebook_new_raw <- bq_table_download(tb)
# 
# saveRDS(facebook_new_raw, file = "./data/Facebook/raw/facebook_new_raw.rds")

sql <- "WITH slice_time_data AS
  (
    SELECT
      *,
      CASE
        WHEN datetime_diff(utc_time, LAG(utc_time, 1) OVER(PARTITION BY postcode, location_type ORDER BY utc_time), SECOND) > 10 THEN utc_time
        WHEN utc_time >= '2019-03-27' AND age_group = '18-65+' THEN utc_time
        WHEN LAG(utc_time, 1) OVER(PARTITION BY postcode, location_type ORDER BY utc_time) IS NULL THEN utc_time
        ELSE NULL
      END AS slice_time
    FROM `twitter-stream-205107.twitter_stream.facebook_updated` 
  )
   
  SELECT
    utc_time,
    postcode,
    location_type,
    age_group,
    gender,
    estimated_mau,
    estimated_dau,
    CONCAT(postcode, '|', location_type, '|', CAST(
      LAST_VALUE(slice_time IGNORE NULLS) OVER(PARTITION BY postcode, location_type ORDER BY utc_time ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW)
    AS STRING)) AS slice_id,
    error_message
  FROM slice_time_data
"

tb <- bq_project_query(project, sql)
facebook_slice_raw <- bq_table_download(tb)
rm(tb, sql)

saveRDS(facebook_slice_raw, file = "./data/Facebook/raw/facebook_slice_raw.rds")
```

```{r fbook_load, eval=FALSE}
# facebook_new_raw <- readRDS("./data/Facebook/raw/facebook_new_raw.rds") 
facebook_slice_raw <- readRDS("./data/Facebook/raw/facebook_slice_raw.rds") 
```

Postcodes:

- **9494** *No usual address* 
- **9797** *Migratory - Offshore - Shipping* 
- **2899** *Norfolk Island*
- **6798** *Christmas Island*
- **6799** *Cocos Islands*

were excluded. 

Data with wrong time stamp were excluded. 

Data with missing `location_type` were also excluded. 

They all returned error messages. 

Data collected before 1st Sept were excluded (sparse & incorrect?).

# Clean dataset

```{r, eval=FALSE}
facebook_slice <- facebook_slice_raw %>% 
  filter(utc_time != anydate("1970-01-01")) %>% 
  filter(utc_time >= anydate("2018-09-01")) %>% 
  filter(postcode != "") %>% 
  filter(postcode != "9494") %>% 
  filter(postcode != "9797") %>% 
  filter(postcode != "2899") %>% 
  filter(postcode != "6798") %>% 
  filter(postcode != "6799") %>% 
  filter(!is.na(location_type)) %>% 
  filter(is.na(error_message)) %>% 
  select(-error_message) %>% 
  group_by(slice_id) %>% 
  # mutate(date_min = min(as.Date(utc_time)), 
  #        date_max = max(as.Date(utc_time))) %>% 
  mutate(date = max(as.Date(utc_time))) %>% 
  ungroup() %>% 
  arrange(slice_id) %>% 
  select(utc_time, date, dplyr::everything())

# length(unique(facebook_new_raw$postcode))
# length(unique(facebook_slice$postcode))
# postcodes <- as.data.frame(sort(table(facebook_slice$postcode), decreasing=T))

saveRDS(facebook_slice, file = "./Data/Facebook/clean/facebook_slice.Rds")
rm(facebook_slice_raw)
```

```{r}
facebook_slice <- readRDS("./data/Facebook/clean/facebook_slice.rds") 
```

```{r, results='asis'}
# skim(facebook_slice)
skim(facebook_slice) %>% skimr::kable()
```

## Results

### Scrap type

Different types of scraps were done for different regions and at different times.

```{r}
ggplot(facebook_slice, aes(x=date)) +
  geom_histogram(binwidth = 1) + 
  facet_wrap(~location_type) + 
  theme_minimal() + 
  xlab("") + ylab("No. of data points")
```

We have three major "batches" of data:

#### Sept 2018

```{r}
facebook_slice_2018_09 <- facebook_slice %>% 
  filter(date <= anydate("2018-11-01")) 
```

Age combined (18-65), men & women combined, collected on ~2weeks for `r comma(length(unique(facebook_slice_2018_09$postcode)))` postocdes.

```{r}
mosaic(~ age_group + gender, data = facebook_slice_2018_09)
```

No. of data points:

```{r}
ggplot(facebook_slice_2018_09, aes(x=date)) +
  geom_histogram(binwidth = 1) + 
  facet_wrap(~location_type) + 
  theme_minimal() + 
  xlab("") + ylab("No. of data points")
```

Example numbers for `4067` postcode

```{r}
facebook_slice_2018_09 %>% 
  filter(postcode == 4067) %>% 
  ggplot(aes(x=utc_time, y=estimated_dau, colour=location_type)) +
  geom_line() +
  theme_minimal() + 
  xlab("") + ylab("Estimated dau")
```


#### Nov 2018

```{r}
facebook_slice_2018_11 <- facebook_slice %>% 
  filter(date >= anydate("2018-10-01") & date <= anydate("2018-12-01"))
```

Age split (18-24 to 60-65; in 5y age bands), men & women separately, collected on ~1day for `r comma(length(unique(facebook_slice_2018_11$postcode)))` postocdes.

```{r}
mosaic(~ age_group + gender, data = facebook_slice_2018_11)
```

No. of data points:

```{r}
ggplot(facebook_slice_2018_11, aes(x=age_group)) +
  geom_histogram(stat = "count") + 
  facet_wrap(location_type~gender, ncol = 2) + 
  theme_minimal() + 
  xlab("") + ylab("No. of data points")
```

#### Dec 2018

```{r}
facebook_slice_2018_12 <- facebook_slice %>% 
  filter(date >= anydate("2018-12-01")) %>% 
  filter(gender == "both") %>%
  select(-gender) %>% 
  droplevels() %>% 
  group_by(postcode, date, location_type) %>% 
  mutate(slice_count = n()) %>% 
  ungroup() %>% 
  filter(slice_count >= 10) %>% 
  group_by(postcode, date, location_type, age_group) %>% 
  arrange(utc_time) %>% 
  filter(row_number()==1) %>% 
  ungroup() %>% 
  group_by(postcode, date, location_type) %>% 
  mutate(slice_count = n()) %>% 
  ungroup() %>% 
  select(utc_time, date, dplyr::everything())

# facebook_slice_2018_12 %>%
#   frq(slice_count)
# 
# facebook_slice_2018_12 %<>% 
#   select(-slice_count)

saveRDS(facebook_slice_2018_12, file = "./Data/Facebook/clean/facebook_slice_2018_12.Rds")
```

Age split (18-24 to 60-65; in 5y age bands), men & women combined, collected from Dec 2018 - till present, `r comma(length(unique(facebook_slice_2018_12$postcode)))` Qld postcodes only.

From `XXXX-XX-XX` data were also collected for `18-65+` category (ie total).

```{r}
mosaic(~ age_group, data = facebook_slice_2018_12)
```

No. of data points:

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  ggplot(aes(x=date)) +
  geom_histogram(binwidth = 1) + 
  facet_wrap(~location_type, ncol = 2) + 
  theme_minimal() + 
  xlab("") + ylab("No. of data points")
```

Example numbers for `4067` postcode:

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == 4067) %>% 
  ggplot(aes(x=utc_time, y=estimated_dau, colour=age_group)) +
  geom_line() +
  theme_minimal() + 
  xlab("") + ylab("Estimated dau")
```

Example numbers for `4067` postcode:

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group == "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == 4067) %>% 
  ggplot(aes(x=utc_time, y=estimated_dau, colour=age_group)) +
  geom_line() +
  theme_minimal() + 
  xlab("") + ylab("Estimated dau")
```

```{r eval=FALSE, include=FALSE}
# Time is UTC - needs to be changed if used!
summary(facebook_slice$utc_time)
table(attr(facebook_slice$utc_time, "tzone"))
# attr(facebook_slice$utc_time, "tzone") <- "Australia/Brisbane"
# summary(facebook_slice$utc_time)
# table(attr(facebook_slice$utc_time, "tzone"))
```

### Dec 2018 dataset - example time series

```{r eval=FALSE, include=FALSE}
fbook_agg_maxmin <- facebook_slice_2018_12 %>% 
  group_by(postcode, location_type, age_group) %>% 
  dplyr::summarise(Min_dau = min(estimated_dau),
                   Max_dau = max(estimated_dau),
                   Dif_dau = Max_dau - Min_dau,
                   Ratio_dau = Max_dau / Min_dau,
                   Min_mau = min(estimated_mau),
                   Max_mau = max(estimated_mau)  ,
                   Dif_dau = Max_mau - Min_mau,
                   Ratio_dau = Max_mau / Min_mau
  ) %>% 
  ungroup()

fbook_agg_median <- facebook_slice_2018_12 %>% 
  group_by(postcode , location_type, age_group) %>% 
  dplyr::summarise(Observations = n(), 
                   # Mean_dau = mean(estimated_dau),
                   Median_dau = median(estimated_dau),
                   # Mean_mau = mean(estimated_mau),
                   Median_mau = median(estimated_mau)
  ) %>% 
  ungroup()

```

#### Postcode **4072** St Lucia:

`location_type` "recently"

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == "4072") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

`location_type` "travelling"

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "travelling") %>% 
  filter(postcode == "4072") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

`location_type` "recently"

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group == "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == "4072") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

`location_type` "travelling"

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group == "18-65+") %>% 
  filter(location_type == "travelling") %>% 
  filter(postcode == "4072") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

#### Postcodes around Townsville 

```{r}
POA_simp %>% 
  filter(POA_CODE16 >= 4810 & POA_CODE16 <= 4818) %>% 
  filter(POA_CODE16 != 4816) %>% 
  qtm(borders = "red", text = "POA_CODE16", fill = NULL)
```

4810 

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == "4810") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

4811 

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == "4811") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

4812 

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == "4812") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```

4814 

```{r}
facebook_slice_2018_12 %>% 
  filter(age_group != "18-65+") %>% 
  filter(location_type == "recently") %>% 
  filter(postcode == "4814") %>% 
  ggplot(aes(x=date, y=estimated_dau, group=age_group, colour=age_group)) +
  geom_point() + 
  theme_minimal() + xlab("") 
```










