---
title: "Airbnb & NVS models"
subtitle: "Experiments with multilevel GLMs"
# author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
mainfont: DejaVu Sans
output: 
  html_document: 
    highlight: pygments
    keep_md: no
    # number_sections: yes
    theme: united
    # toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, warning = FALSE, 
                      fig.width = 8, fig.height = 5,
                      out.width = "800px", out.height = "500px",
                      dpi = 300)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

options(scipen=999)
set.seed(12345)

library(pacman) 
p_load(tidyverse, lubridate, sjPlot,
       lme4, ggeffects, performance)

```

<!-- ------------------------------------------------------------ --> 

# Data

```{r}
data <- readRDS("data/model_input/tempo_sa3_18_month_long.Rds")

glimpse(data)

# length(unique(data$SA3_CODE16))
# length(unique(data$SA3_NAME16))
```

## Airbnb

`Illawarra Catchment Reserve` has no Airbnb data. TS filled with zeros.

```{r}
data <- data %>% 
  mutate(Airbnb_visitor_nights = ifelse(is.na(Airbnb_visitor_nights), 
                                        0, Airbnb_visitor_nights))
```

### Individual time trajectories

```{r}
summary(data$Airbnb_visitor_nights)

data %>% 
  ggplot(aes(x = date, y = Airbnb_visitor_nights/100000)) + 
  geom_line(aes(group = SA3_CODE16), 
            alpha = 0.15, color = "darkorchid4") + 
  theme_minimal()
```


### Freq of values

```{r}
data %>%  
  ggplot(aes(x = Airbnb_visitor_nights)) + 
  geom_histogram(binwidth = 1000) + 
  theme_minimal()

```

### Totals by region

```{r}
aggregated <- data %>% 
  group_by(SA3_CODE16, SA3_NAME16) %>% 
  summarise(NVS_dest_nights_year = as.integer(round(sum(NVS_dest_nights))), 
            Airbnb_visitor_nights_year = as.integer(round(sum(Airbnb_visitor_nights)))) %>%   
  ungroup()

summary(aggregated$Airbnb_visitor_nights_year)

aggregated %>%  
  ggplot(aes(x = Airbnb_visitor_nights_year/100000)) + 
  geom_histogram(binwidth = 1) + 
  theme_minimal()

```

## NVS

### Individual time trajectories

```{r}
summary(data$NVS_dest_nights)

data %>% 
  ggplot(aes(x = date, y = NVS_dest_nights/100000)) + 
  geom_line(aes(group = SA3_CODE16), 
            alpha = 0.15, color = "darkorchid4") + 
  theme_minimal()

```

### Totals by region

```{r}
summary(aggregated$NVS_dest_nights_year)

aggregated %>% 
  ggplot(aes(x = NVS_dest_nights_year/100000)) + 
  geom_histogram(binwidth = 1) + 
  theme_minimal()

```


## Relationship

### Yearly 

```{r}
aggregated %>% 
  ggplot(aes(x = NVS_dest_nights_year/100000)) + 
  geom_point(aes(y = Airbnb_visitor_nights_year/100000), 
             alpha = 0.5) + 
  geom_line(aes(y = NVS_dest_nights_year/100000), 
            color = "lightblue") +
  theme_minimal()
```

Only few places where Airbnb is equal or more NVS.

```{r}
aggregated %>% 
  filter(Airbnb_visitor_nights_year >= NVS_dest_nights_year) %>% 
  dplyr::select(SA3_NAME16, Airbnb_visitor_nights_year, NVS_dest_nights_year)

```

### Monthly

```{r}
data %>% 
  group_by(SA3_CODE16) %>% 
  ggplot(aes(x = NVS_dest_nights)) + 
  geom_point(aes(y = Airbnb_visitor_nights), 
             alpha = 0.5) + 
  geom_line(aes(y = NVS_dest_nights), 
            color = "lightblue") + 
  facet_wrap(~date) +
  theme_minimal()
```

More places where Airbnb is equal or more NVS.

```{r}
data %>% 
  group_by(SA3_CODE16) %>% 
  filter(Airbnb_visitor_nights >= NVS_dest_nights) %>% 
  dplyr::select(SA3_NAME16, Airbnb_visitor_nights, NVS_dest_nights)

```

# Poisson models using `lme4`

```{r}
data <- data %>% 
  left_join(aggregated)

data <- data %>% 
  group_by(SA3_CODE16) %>% 
  mutate(time = row_number()) %>% 
  ungroup()
```

## random region, no time

```{r}
m1 <- glmer(Airbnb_visitor_nights ~ 1 + 
              offset(log(NVS_dest_nights_year)) +
              (1 | SA3_NAME16),
            data = data, 
            family = poisson(link = "log"))

summary(m1)
# confint(m1)
```

## random region + time

```{r}
m2 <- glmer(Airbnb_visitor_nights ~ 1 + 
              time +
              offset(log(NVS_dest_nights_year)) +
              (1 | SA3_NAME16),
            data = data, 
            family = poisson(link = "log"))

summary(m2)
# confint(m2)
```

## random region + random time

```{r}
m3 <- glmer(Airbnb_visitor_nights ~ 1 + 
              time +
              offset(log(NVS_dest_nights_year)) +
              (1 + time | SA3_NAME16),
            data = data, 
            family = poisson(link = "log"))

summary(m3)
# confint(m3)
```

## Comparison

```{r}
tab_model(m1, m2, m3)

performance::compare_performance(m1, m2, m3)
# performance::model_performance(m3)

performance::check_overdispersion(m3)
performance::check_singularity(m3)
performance::check_model(m3, panel = TRUE)
performance::r2(m3) 
performance::icc(m3) 
insight::get_variance(m3)

```

## Best model (AIC/BIC)

```{r}
summary(m3)

# anova(m1, m3)
# anova(m2, m3)
# anova(m3, m3)

# plot_model(m3, type = "est")
# plot_model(m3, type = "slope")
# plot_model(m3, type = "resid")
# plot_model(m3, type = "diag")

data$m3_fit_poi <- fitted(m3)

ggplot(data, aes(x = Airbnb_visitor_nights)) +
  geom_point(aes(y = m3_fit_poi),
             alpha = 0.3) +
  geom_line(aes(y = Airbnb_visitor_nights), col = "darkorchid4") +
  theme_minimal() + xlab("Airbnb_visitor_nights") + ylab("Predicted")

ggplot(data, aes(x = Airbnb_visitor_nights)) +
  geom_point(aes(y = m3_fit_poi),
             alpha = 0.3) +
  geom_line(aes(y = Airbnb_visitor_nights), col = "darkorchid4") +
  facet_wrap(~date) +
  theme_minimal() + xlab("Airbnb_visitor_nights") + ylab("Predicted")

# plot(ggpredict(m3, "time"))

```

## Example areas

```{r}
regions <- readRDS("./data/regions/regions.rds") %>% 
  dplyr::select(sa3, region_types)  %>% 
  rename(SA3_NAME16 = sa3)

regions
```

```{r}
data %>% 
  right_join(regions) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_poi, group = SA3_NAME16, col = region_types)) +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  theme_minimal() + 
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

```{r}
data %>% 
  right_join(regions) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_poi, group = SA3_NAME16, col = region_types)) +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  geom_point(aes(y = NVS_dest_nights),
             col = "#E7B800", alpha = 0.33) +
  theme_minimal() + 
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

## Outliers

### Absolute

```{r}
data$diff_abs <- data$Airbnb_visitor_nights - data$m3_fit_poi

data %>% 
  arrange(desc(diff_abs)) %>% 
  dplyr::select(SA3_NAME16, date, diff_abs) %>% 
  print(n = 20)

data %>% 
  arrange(diff_abs) %>% 
  dplyr::select(SA3_NAME16, date, diff_abs) %>% 
  print(n = 20)
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("Mornington Peninsula", 
             "Surf Coast - Bellarine Peninsula",
             "Shoalhaven",
             "Melbourne City"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_poi, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("Mornington Peninsula", 
             "Surf Coast - Bellarine Peninsula",
             "Shoalhaven",
             "Melbourne City"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_poi, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  geom_point(aes(y = NVS_dest_nights),
             col = "#E7B800", alpha = 0.33) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

### Relative

```{r}
data$diff_rel <- (data$Airbnb_visitor_nights - data$m3_fit_poi)/data$Airbnb_visitor_nights

data %>% 
  filter(Airbnb_visitor_nights > 0) %>%  
  arrange(desc(diff_rel)) %>% 
  dplyr::select(SA3_NAME16, date, diff_rel) %>% 
  print(n = 20)

data %>% 
  filter(Airbnb_visitor_nights > 0) %>%  
  arrange(diff_rel) %>% 
  dplyr::select(SA3_NAME16, date, diff_rel) %>% 
  print(n = 20)
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("Port Stephens", 
             "Barkly",
             "Central Highlands (Qld)",
             "Canberra East"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_poi, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("Port Stephens", 
             "Barkly",
             "Central Highlands (Qld)",
             "Canberra East"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_poi, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  geom_point(aes(y = NVS_dest_nights),
             col = "#E7B800", alpha = 0.33) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

# NB models using `lme4`

## Distributions

```{r}
p_load(fitdistrplus)

poi <- fitdist(data$Airbnb_visitor_nights, "pois")
nb <- fitdist(data$Airbnb_visitor_nights, "nbinom")

par(mfrow = c(1,2))
denscomp(list(poi, nb), 
         legendtext = c("Poisson", "NB"), fitlty = 1)
cdfcomp(list(poi, nb), 
        legendtext = c("Poisson", "NB"), fitlty = 1)

gofstat(list(poi, nb), 
        fitnames = c("Poisson", "negative binomial"))
```

## random region, no time

```{r}
m1 <- glmer.nb(Airbnb_visitor_nights ~ 1 + 
              offset(log(NVS_dest_nights_year)) +
              (1 | SA3_NAME16),
            data = data)

summary(m1)
# confint(m1)
```

## random region + time

```{r}
m2 <- glmer.nb(Airbnb_visitor_nights ~ 1 + 
              time +
              offset(log(NVS_dest_nights_year)) +
              (1 | SA3_NAME16),
            data = data)

summary(m2)
warnings()
# confint(m2)
```

## random region + random time

```{r}
m3 <- glmer.nb(Airbnb_visitor_nights ~ 1 + 
              time +
              offset(log(NVS_dest_nights_year)) +
              (1 + time | SA3_NAME16),
            data = data)

summary(m3)
# confint(m3)
```

## Comparison

```{r}
tab_model(m1, m2, m3,
          show.aic = TRUE, show.dev = TRUE)

performance::compare_performance(m1, m2, m3)
# performance::model_performance(m3)

performance::check_singularity(m3)
performance::check_model(m3, panel = TRUE)
performance::r2(m3) 
performance::icc(m3) 
insight::get_variance(m3)

```

## Best model (AIC/BIC)

```{r}
summary(m3)

# anova(m1, m3)
# anova(m2, m3)
# anova(m3, m3) 

# plot_model(m3, type = "est")
# plot_model(m3, type = "slope")
# plot_model(m3, type = "resid")
# plot_model(m3, type = "diag")

data$m3_fit_nb <- fitted(m3)

ggplot(data, aes(x = Airbnb_visitor_nights)) +
  geom_point(aes(y = m3_fit_nb),
             alpha = 0.3) +
  geom_line(aes(y = Airbnb_visitor_nights), col = "darkorchid4") +
  theme_minimal() + xlab("Airbnb_visitor_nights") + ylab("Predicted")

ggplot(data, aes(x = Airbnb_visitor_nights)) +
  geom_point(aes(y = m3_fit_nb),
             alpha = 0.3) +
  geom_line(aes(y = Airbnb_visitor_nights), col = "darkorchid4") +
  facet_wrap(~date) +
  theme_minimal() + xlab("Airbnb_visitor_nights") + ylab("Predicted")

# plot(ggpredict(m3, "time"))

```

## Example areas

```{r}
data %>% 
  right_join(regions) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_nb, group = SA3_NAME16, col = region_types)) +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  theme_minimal() + 
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

```{r}
data %>% 
  right_join(regions) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_nb, group = SA3_NAME16, col = region_types)) +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  geom_point(aes(y = NVS_dest_nights),
             col = "#E7B800", alpha = 0.33) +
  theme_minimal() + 
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

## Outliers

### Absolute

```{r}
data$diff_abs <- data$Airbnb_visitor_nights - data$m3_fit_nb

data %>% 
  arrange(desc(diff_abs)) %>% 
  dplyr::select(SA3_NAME16, date, diff_abs) %>% 
  print(n = 20)

data %>% 
  arrange(diff_abs) %>% 
  dplyr::select(SA3_NAME16, date, diff_abs) %>% 
  print(n = 20)
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("Mornington Peninsula", 
             "Surf Coast - Bellarine Peninsula",
             "Shoalhaven",
             "Melbourne City"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_nb, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("Mornington Peninsula", 
             "Surf Coast - Bellarine Peninsula",
             "Shoalhaven",
             "Melbourne City"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_nb, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  geom_point(aes(y = NVS_dest_nights),
             col = "#E7B800", alpha = 0.33) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

### Relative

```{r}
data$diff_rel <- (data$Airbnb_visitor_nights - data$m3_fit_nb)/data$Airbnb_visitor_nights

data %>% 
  filter(Airbnb_visitor_nights > 0) %>%  
  arrange(desc(diff_rel)) %>% 
  dplyr::select(SA3_NAME16, date, diff_rel) %>% 
  print(n = 20)

data %>% 
  filter(Airbnb_visitor_nights > 0) %>%  
  arrange(diff_rel) %>% 
  dplyr::select(SA3_NAME16, date, diff_rel) %>% 
  print(n = 20)
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("East Arnhem", 
             "Barkly",
             "Central Highlands (Qld)",
             "Canberra East"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_nb, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```

```{r}
data %>% 
  filter(SA3_NAME16 %in% 
           c("East Arnhem", 
             "Barkly",
             "Central Highlands (Qld)",
             "Canberra East"
           )) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = m3_fit_nb, group = SA3_NAME16), col = "darkorchid4") +
  geom_point(aes(y = Airbnb_visitor_nights), 
             col = "#00AFBB", alpha = 0.3) +
  geom_point(aes(y = NVS_dest_nights),
             col = "#E7B800", alpha = 0.33) +
  scale_x_continuous(limits = c(0, 12), breaks = c(2:10)) +
  theme_minimal() + 
  xlab("") + ylab("Observed/fitted") + 
  facet_wrap(~SA3_NAME16, scales = "free_y")
```