---
title: "Sensing ERP - EDA 02"
author: "Radek Panczak et al."
date: "`r format(Sys.time(), '%d %B, %Y')`"
mainfont: DejaVu Sans
output: 
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    theme: united 
    highlight: pygments 
---

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=8, fig.height=6, dpi=300, out.width="800px", out.height="600px")

knitr::opts_knit$set(root.dir = 'R:/ABSLINKAGE-A1760/ABS Linkage')

options(scipen=999)
set.seed(12345)

library(pacman) 
p_load(tidyverse, readxl, magrittr, here, janitor,
       skimr, sjmisc, naniar, kableExtra, statar, scales, anytime,
       sf, tmap, tmaptools)

tmap_mode("view") # makes map interactive

```

# Twitter

## Data

```{r}
tweets_geo2_inside <- readRDS(file = "R:/POP247-Q0786/data/Twitter/clean/tweets_geo2_inside.rds")
```

**`r nrow(tweets_geo2_inside)`** georeferenced (point level) tweets from `tweets_geo2_inside` dataset was used.

```{r}
tweets_geo2_proc <- tweets_geo2_inside %>% 
  select(-source, -place_id, -place_country, -place_bounding_box, 
         -user_utc_offset, -user_time_zone,
         -user_followers_count, -user_friends_count, -user_favourites_count, -user_created_at,
         -in_reply_to_status_id_str, -in_reply_to_user_id_str,
         -geo_lat, -geo_lon, -orig_lat, -orig_lon) %>% 
  mutate(date = as.Date(created_at)) %>% 
  filter(date >= anydate("2018-12-01") & date <= anydate("2019-02-28"))
```

**`r nrow(tweets_geo2_proc)`** tweets from Dec '19 - Feb '19 were included. They were linked to postcodes and unique number of users per month was calculated.

```{r}
tweets_geo2_join <- tweets_geo2_proc %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  st_transform(3112) %>% 
  st_join(POA, join = st_intersects) %>% 
  filter(!is.na(POA_CODE16))
```

```{r}
tweets_geo2_agg <- tweets_geo2_join %>% 
  st_drop_geometry() %>% 
  mutate(month = format(date, "%y-%m")) %>% 
  group_by(POA_CODE16, month) %>% 
  summarise(unique_users = n_distinct(UID))
```

## Results 

### Number of distinct users per month per postcode:

```{r}
tweets_geo2_agg %>% 
  ggplot(aes(x=month, y=unique_users)) +
  geom_boxplot() + scale_y_log10()
```

Seems like Feb is affected by data collection outage?

### Spatial view of Dec

```{r}
tweets_geo2_poa <- POA_centr %>% 
  left_join(tweets_geo2_agg) %>% 
  filter(month == "18-12" | month == "19-01")

tweets_geo2_poa %>% 
  filter(month == "18-12") %>% 
  tm_shape() +
  tm_bubbles("unique_users", scale=.5) + 
  tm_facets(by = "month")
```

### Differences Jan - Dec

```{r}
tweets_geo2_agg %>% 
  filter(month == "18-12" | month == "19-01") %>% 
  spread(month, unique_users) %>% 
  mutate(dif = `19-01` - `18-12`) %>% 
  ggplot(aes(dif))+
  geom_histogram()
```

### Ratio to Census pops J

Unique users / Pop 18+ multiplied by 1000

```{r}
POA_twitter <- POA %>% 
  left_join(pop_POA) %>% 
  left_join(tweets_geo2_agg) %>% 
  mutate(Twitter_ratio = (unique_users / pop_cen_18plus) *1000)
```

#### POA without pops, with Tweets

```{r}
POA_twitter %>% 
  filter(month == "18-12" | month == "19-01") %>% 
  filter(Twitter_ratio == Inf) %>% 
  tm_shape() +
  tm_dots(col = "red") + 
  tm_facets(by = "month", sync = TRUE)
```

Seems to pick up some airports, campuses & shopping centres.

#### POA with high ratio

```{r}
POA_twitter %>% 
  filter(month == "18-12" | month == "19-01") %>% 
  filter(Twitter_ratio >= 100 & Twitter_ratio < Inf) %>% 
  tm_shape() +
  tm_dots(col = "red") + 
  tm_facets(by = "month", sync = TRUE)
```

Seems to pick up industrial areas (Malaga in Perth), remorte areas & parks.

#### Ratio distribution

Excluding two categories mapped above

```{r}
POA_twitter %>% 
  filter(month == "18-12" | month == "19-01") %>% 
  filter(Twitter_ratio < 100) %>% 
  filter(month == "18-12") %>% 
  ggplot(aes(Twitter_ratio))+
  geom_histogram(binwidth = 1)
```

```{r}
POA_twitter %>% 
  filter(month == "18-12" | month == "19-01") %>% 
  filter(Twitter_ratio < 100) %>% 
  # filter(month == "18-12") %>%
  tm_shape() +
  tm_fill(col = "Twitter_ratio", style = "fixed", palette = "div",
          breaks = c(0, 1, 2, 5, 10, 20, 30)) + 
  tm_facets(by = "month", sync = TRUE)
```

# Airbnb

## Data

```{r}
POA_2016_join_new <- readRDS(file = "R:/AIRBNB-Q0931/data/airdna/clean/POA_2016_join_new.rds")

monthly_new_select <- readRDS("R:/AIRBNB-Q0931/data/airdna/clean/monthly_new_select.Rds") %>% 
  filter(reporting_month >= as.Date("2018-12-01")) %>% 
  filter(reporting_month <= as.Date("2019-01-01")) %>% 
  filter(number_of_reservations > 0) %>% 
  left_join(POA_2016_join_new)
```

`r nrow(monthly_new_select)` records of `r length(unique(monthly_new_select$property_id))` properties that have at least one bookin in Dec and Jan were used. `r sum(monthly_new_select$number_of_reservations, na.rm = TRUE)` bookings were agggregated by postcode.

```{r}
temp <- monthly_new_select %>% 
  group_by(POA_CODE16, reporting_month) %>% 
  summarise(number_of_reservations = sum(number_of_reservations, na.rm = TRUE)) 

POA_airbnb <- POA %>% 
  left_join(pop_POA) %>% 
  left_join(temp)%>% 
  mutate(Airbnb_ratio = (number_of_reservations / pop_cen_18plus))

rm(monthly_new_select, POA_2016_join_new, temp)
gc()
```


## Results 

### Number of reservations per month per postcode:

```{r}
POA_airbnb %>% 
  filter(!is.na(reporting_month)) %>% 
  ggplot(aes(x=as.factor(reporting_month), y=number_of_reservations)) +
  geom_boxplot() + scale_y_log10()
```

### Spatial view of Dec

```{r}
POA_airbnb %>% 
  filter(reporting_month == as.Date("2018-12-01")) %>% 
  tm_shape() +
  tm_bubbles("number_of_reservations", scale=.5)
```

### Ratio to Census pops J

Number of reservations / Pop 18+

#### POA without pops, with bookings

```{r}
POA_airbnb %>% 
  filter(Airbnb_ratio == Inf) %>% 
  tm_shape() +
  tm_dots(col = "red") + 
  tm_facets(by = "reporting_month", sync = TRUE)
```


#### POA with high ratio

Ratio equal or more than 1 ie. same amount of reservations as pops

```{r}
POA_airbnb %>% 
  filter(Airbnb_ratio >= 1 & Airbnb_ratio < Inf) %>% 
  tm_shape() +
  tm_dots(col = "red") + 
  tm_facets(by = "reporting_month", sync = TRUE)
```

#### Ratio distribution

Excluding first category above

```{r}
POA_airbnb %>% 
  filter(reporting_month == as.Date("2018-12-01")) %>% 
  filter(Airbnb_ratio < Inf) %>% 
  ggplot(aes(Airbnb_ratio))+
  geom_histogram(binwidth = 0.1)
```

```{r}
POA_airbnb %>% 
  filter(Airbnb_ratio < Inf) %>% 
  tm_shape() +
  tm_fill(col = "Airbnb_ratio", style = "fixed", palette = "div",
          breaks = c(0, 0.1, 0.5, 1, 2)) + 
  tm_facets(by = "reporting_month", sync = TRUE)
```

